Automated verification of electrical systems: literature review and case studies of stand-alone solar photovoltaics\thanks{Supported by Newton Fund (ref. 261881580) and FAPEAM (Amazonas State Foundation for Research Support, calls 009/2017 and PROTI Pesquisa 2018).}

### 2019-01-18 Tex/CAV2019/Trindade_Cordeiro_CAV2019.tex
CAV2019 paper

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
%\else
  % normal IEEE
%  \usepackage{cite}
%\fi
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
\usepackage{graphicx}
\graphicspath{ {./images/} }
%
%\ifCLASSINFOpdf
  %\usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
%  \DeclareGraphicsExtensions{.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
%\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithm,algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

%Alessandro input: makecell package to permite line break inside tables
\usepackage{makecell}
\usepackage{xcolor}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%\renewcommand{\baselinestretch}{0.975}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
%\title{Building a framework to perform automated verification of solar photovoltaic off-grid systems}
\title{Automated Verification of Stand-alone Solar Photovoltaic Systems}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Alessandro~Trindade and Lucas~Cordeiro% <-this % stops a space
\IEEEcompsocitemizethanks{
\IEEEcompsocthanksitem{A. Trindade is with the Department of Electricity, Federal University of Amazonas, Manaus, Brazil, e-mail: alessandrotrindade@ufam.edu.br.}% <-this % stops a space
\IEEEcompsocthanksitem{L. Cordeiro is with School of Computer Science, The University of Manchester, UK, e-mail: lucas.cordeiro@manchester.ac.uk.}}% <-this % stops a space
\thanks{The authors would like to thank Newton Fund (ref. 261881580) and FAPEAM (Amazonas State Foundation for Research Support, call 009/2017), for the financial support.}
\thanks{Manuscript received November 23, 2018; revised XXXXX 11, 2019.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Computers,~Vol.~14, No.~8, August~2015}%
{Trindade \MakeLowercase{\textit{et al.}}: Automated Verification of Stand-Alone Solar Photovoltaic Systems}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. Automated verification using model checking has proven to be an effective technique to program verification. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system including solar panel, charge controller, battery, inverter, and electric load. Case studies, from real photovoltaic systems deployed in five different sites, ranging from 700W to 1,200W, were used to evaluate this proposed approach and to compare that with specialized simulation tools. Data from practical applications show the effectiveness of our approach, where specific conditions that lead to failures in a photovoltaic solar system are only detected by our automated verification method.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Formal verification, model checking, photovoltaic power systems, power system modeling, solar power generation.
\end{IEEEkeywords}}

% make the title area
\maketitle

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{here} are presently 1.3 billion people with no access to electricity worldwide~\cite{Coelho}. 
%The definition of energy access usually includes both electricity access and access to modern fuels for cooking and heating, in order to replace traditional biomass. 
%Increased access to energy allows economic growth and poverty alleviation \cite{Karekesi}. 
%
%Among the options of renewable sources, there are hydro, wind, and solar photovoltaic (PV). 
Only a niche market a few years ago, solar photovoltaic (PV) systems are now becoming a mainstream electricity provider, with an increase of approximately 50\% from 2016 to 2017 in terms of new installations of PV~\cite{EPIA}.

In order to simulate or evaluate a PV system there are a myriad of specialized tools available in the market, such as RETScreen, HOMER, PVWatts, SAM, and Hybrid2 \cite{Pradhan,Swarnkar,NRELDobos,NRELBlair,Mills}; and even general purpose simulation tools such as PSpice, Saber, or MATLAB/Simulink package \cite{Gow1999,Benatiallah2017}.
%
%However, to the best of our knowledge, this work can be the first work to perform automated verification of a stand-alone solar PV project solution.  
%
%According to \cite{SEIA} and \cite{Chauhan}, solar is the most abundant source of renewable energy on earth. In SPV system, solar radiations are captured from the sun and turned into electricity using SPV cells made of silicon and other materials. SPV...
%
%In order to address different aspects of a SPV project, there is a myriad of public domain and commercial software available. According to \cite{Brooks}, the capabilities of these tools range from simple solar resource and energy production estimative, to site survey and system design tools, to complex financial analysis software (with optimization). Some tools also provide support to rebate programs applications and tax incentives (specific to each country or region), while other programs and worksheets focus on the technical aspects of system sizing and design.  
%
%Manufactures and integrators have yet their proprietary software to perform various system sizing, with the drawback of include just their own products among the possibilities of choice, what restrict the solution. 
%
%PV systems can be classified into grid-connected and stand-alone systems. At this proposal, only off-grid or stand-alone systems will be considered. The utilization of solar energy in off-grid mode has the potential to meet the energy need for remote rural areas of developing countries.  
However, those tools are based on running experiments in simulation models. Simulation has the advantage of being cheap (if compared to test in real systems) and can be employed before the system design is concluded but it has the drawback of an incomplete coverage since the verification of all possible combinations and potential failures of a system is unfeasible~\cite{ClarkeHV18}.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{makecell}
\usepackage{algorithm,algorithmic}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Automated verification of electrical systems: literature review and case studies of stand-alone solar photovoltaics\thanks{Supported by Newton Fund (ref. 261881580) and FAPEAM
(Amazonas State Foundation for Research Support, calls 009/2017 and PROTI Pesquisa 2018).}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alessandro Trindade\inst{1}\orcidID{0000-0001-8262-2919} \and
Lucas Cordeiro\inst{2}\orcidID{0000-0002-6235-4272}}
%
\authorrunning{A. Trindade and L. Cordeiro}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Amazonas, Av. Gen. Rodrigo Octavio, 6200, Coroado I, 69077-000 Manaus AM Brazil \email{alessandrotrindade@ufam.edu.br} \and
The University of Manchester, Kilburn Building, Oxford Road, Manchester M13 UK
\email{lucas.cordeiro@manchester.ac.uk}}
% \and ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Around 1 billion of people do not have electrical energy nowadays on Earth. Energy poverty alleviation has become an important political issue. Several initiatives and policies have been proposed to deal with poor access to sources of energy in many developing countries. Particular attention is given to use renewables and stand-alone solar photovoltaic systems. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. The use of automated verification in electrical systems is a fresh subject with few issued literature. This paper do a state-of-art review about the use of automated verification on electrical systems and perform a comparative of tolls applied to five stand-alone solar photovoltaic systems. Information about the use of a commercial simulator and monitoring data from the deployed systems are used to validate the research. Data from experimental results shows that only some flaws can be detected by automated verification and that ESBMC verification engine obtain the best results, regardless using a high-end or low-end setup.

\keywords{Automated verification  \and Model Checking \and Electrical systems \and Solar photovoltaic systems.}
\end{abstract}
%
%
%
\section{Introduction}
%%%%%In 2015, 193 Member States of the United Nations agreed as part of the Sustainable Development Goals (SDG) with a specific target to 'ensure access to affordable, reliable and modern energy for all by 2030' – universal access to electricity and clean cooking~\cite{IEAeao2017}. Therefore electricity has a important role, based on two main issues: %that are on the agenda of all current rulers and society in general, regardless of whether it is a developed, emerging or developing country: 
%%%%%the impact on the global climate change and poverty alleviation.
%
%%%%%Related to climate change, the 23rd session of the Conference of the Parties (COP23) to the United Nations Framework Convention on Climate Change (UNFCCC) in 2017 was the second session after the Paris Agreement, which aims to keep the maximum global average temperature rise as close as possible to 1.5 $^{o}$C. %During COP23, with respect to the future market development of renewable power and solar PV (photovoltaics) in particular, it was created an alliance of nations and states committed to moving the world from burning coal to cleaner power sources was a promising sign at COP23 \cite{Jager-Waldau}. 
%%%%%Decarbonization of our energy supply is an important component to achieve the targets, because 65.3\% of the worlds current CO\textsubscript{2} emissions (from energy supple) are due to burning fossil fuels according with 2016 data: 81.1\% of our total primary energy supply depended on burning fossil fuels, namely 27.1\% coal, 31.9\% oil and 22.1\% natural gas~\cite{IEAwes2018}. In terms of final energy consumption electricity only accounted for 18.6\%, but was responsible for 32\% of the total CO\textsubscript{2} emissions~\cite{IEAwes2018}. Photovoltaics (PV) is one of the key technology options for implementing the shift to a decarbonized energy supply, because it is abundant, cannot be monopolized  and can be deployed in a modular way almost everywhere on Earth.
%
%Related to poverty alleviation, 
%
%In reality, the biggest barrier to universal energy access is not the capacity to generate electricity, but rather the ability to get energy to those who need it most. The concept of energy poverty and access tends to focus on very small, incremental shifts in the delivery of energy to poor people. The total energy demand of these shifts constitutes only a small fraction of the ‘gap’ between electricity supply in much of Africa versus more industrialised countries, or the gap between the region’s current and projected electricity supply. https://policy-practice.oxfamamerica.org/static/media/files/FINAL_speakingpowertotruth_SH.pdf
%
%%%%%Poverty, in other hand, needs to be addressed through socio-economic development. Poverty %is conceptualized in material terms as not having access to adequate levels of food, water, clothing, shelter, sanitation, health care and education. This 
%%%%%can be translated into people having insufficient income. 
%
For numerous people around the world, a better life means getting access to basic needs such as food, health services, housing and clean water. None of these basic needs can be provided without energy. Energy is one of the most essential inputs into sustaining people‘s livelihoods. Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}.

Progress has been made worldwide. The number of people without electricity access fell below 1 billion for the first time in 2017~\cite{IEAweo2018}. To provide universal electricity for all, decentralized systems, led by solar PV in off-grid and mini-grid systems, will be the least-cost solution for three-quarters of the additional connections needed, but that grid extension will still have a role to play, especially in urban areas~\cite{IEAweo2018}.

%With worldwide over 400GW cumulative installed photovoltaic electricity generation capacity installed by the end of 2017, photovoltaics still is a small contributor to the electricity supply with a little of 2\%, but its importance for our future energy mix is now acknowledged and with the fastest growth (32\% whem compared with 2016 numbers)~\cite{Jager-Waldau}.
%
In order to simulate or evaluate a PV system there are a myriad of specialized tools available in the market, such as RETScreen, HOMER, PVWatts, SAM, and Hybrid2 \cite{Pradhan,Swarnkar,NRELDobos,NRELBlair,Mills}; and even general purpose simulation tools such as PSpice, Saber, or MATLAB/Simulink package \cite{Gow1999,Benatiallah2017}. However, those tools are based on running experiments in simulation models. Simulation %has the advantage of being cheap (if compared to test in real systems) and can be 
is employed before the system design is concluded but it has the drawback of an incomplete coverage since the verification of all possible combinations and potential failures of a system is unfeasible~\cite{ClarkeHV18}.

Formal methods based on model checking offer a great potential to obtain a more effective and faster verification in the design process~\cite{ClarkeHV18}. Any kind of system can be specified as computer programs using mathematical logic, which constitutes the intended (correct) behavior; then, one can try to give a formal proof or otherwise establish that the program meets its specification.

In this study, an updated review about the use of verification tools in electrical systems is performed, and is presented five case studies, where different model checking tools are compared in terms of performance and reliability when validating stand-alone solar PV systems. In order to perform the comparation and case studies, a mathematical model of each component of a stand-alone PV system, as panel solar, charge controller, batteries, inverter, and electrical load is used. The project requirements, as battery autonomy and power demand, besides weather conditions, as solar irradiance and ambient temperature, are translated as part of the computer program and automatically verified during the formal process. The model checking tools report in which conditions a system does not meet the user requirements. A key benefit to this approach is that it helps in the detection of flaws in the design phase of system development, thereby considerably improving system reliability~\cite{Akram2018}. The case studies are not theoretical and were deployed in 2018, with feedback information collected from monitoring systems and through interview with the dwellers. 
%The implementation of the proposed tool is carried out by means of the efficient SMT-based bounded model checker (ESBMC)~\cite{esbmc2018}.
\section{State-of-art in automated verification of Electrical systems}
This section summarizes, in a chronological way, the research done by previous studies related to the subject of this research.

The conversion of traditional power grid into a smart grid, a fundamental example of a cyber-physical system, raises a number of issues that require novel methods and applications. In 2012, it was considered a specific Chinese Smart Grid implementation as a case study and address the verification problem for performance and energy consumption~\cite{Yukseletall2012}. It was employed the stochastic model checking approach and present a modelling and analysis study using PRISM model checker. The focus was how Cyber-physical systems integrate information and communication technology functions to the physical elements of a system for monitoring and controlling purposes.  In this context, it was applied automated verification of certain quantitative properties of the system. %, with no interest in power generation or even solar PV systems.

%In 2014, there were a research showing that power systems model information exchange and the simulation of user defined models is a challenge issue for the transmission system operator. The reason: the use of different simulation software with specific data format and models strongly coupled to the simulation software numerical routines. The researchers proposed the use of Open Source software (Modelica, JAVA and Python) for simulation and modelling as a solution to decouple these models. The work described the design of a simulation tool following the Free/Libre Open Source Software philosophy to simulate equation-based models and implement a standard data format for model information exchange~\cite{Gomez2014}. The aim was the use of Modelica to define standards and industry models, besides the interest of components in modules. This work did not reached the automated verification, it is just simulation, however created the base to future work of formal methods on electrical systems.
%
In 2015, it was showed that related to power system protection schemes, the approach of using Monte-Carlo simulation and manual interpretation of results had the limitation of the incomplete coverage of all possible operating conditions due to the time complexity of simulation~\cite{Sengupta2015}. It was proposed an automated simulation-based verification technique to verify the correctness of protection settings efficiently using hybrid automata-temporal-logic framework. The initial focus were relay operations and test-case generation to ensure early detection of design errors. %The study was limited to power system protection and did not deal with electricity generation or even solar PV systems.

Other work issued in 2015 was a framework named Modana to achieve an integrated process from modeling with SysML/MARTE to analysis with Statistical Model Checking (SMC) for Cyber-Physical Systems (CPSs) in terms of Non Functional Properties (NFP) such as time, energy, etc~\cite{Cheng2015}. To demonstrate the capability of Modana framework, the authors modelled energy-aware buildings as a case study, and discussed the analysis on energy consumption in different scenarios. The focus is in smart buildings and HVAC (Heating, ventilation, and air conditioning) systems. %The research did no deal with solar PV systems.

Back in 2017, it was suggested the use of formal methods to verify and certifiably control the behavior of computational devices interacting over a shared and smart infrastructure~\cite{Abate2017}. It was shown that formal techniques can provide compelling solutions not only when safety-critical goals are the target, but also to tackle verification and synthesis problems on populations of such devices. This work argued that alternative solutions based on classical analytical techniques or on approximate computations are prone to errors with global repercussions. Two different areas were targeted: thermal loads and electricity-generating devices interacting over a smart energy network or over a local power grid. The researcher discussed the aggregation of large populations of thermostatically-controlled loads and of photovoltaic panels, and the corresponding problems of energy management in smart buildings, of demand-response on smart grids, and respectively of frequency stabilization and grid robustness. The focus is in controlling the behavior of the components, verifying the smart grid, dealing with a 'system of systems' and with 'internet of things'. The author uses approximate model checking of stochastic and hybrid models. % Therefore, the focus is not in electricity generation or even the components of a stand-alone solar PV system.

In 2018, is was proposed a methodology for the verification of the Cyber Physical Energy Systems (CPES), with the methodology applied to PV panels and its distributed power point tracking~\cite{Driouich2018}. The approach relied on representing the unpredictable behavior of the environment in order to cover all feasible possible scenarios. Processed by JModelica, the simulation results covered the system's complete dynamic behavior. It was evident at this work the time consuming issue, with almost three days of computer effort to verify the design space of one hour from PV panels behavior. %The work did not include the other components of a stand-alone solar PV system.

Another work issued in 2018 was the approach to modeling smart grid components using a formal specification. It was used a state-based formal specification language namely Z %(pronounced as `Zed') 
and it was demonstrated the application of Z on four smart grid components~\cite{Akram2018}. The presented formal specification can be considered as first steps towards modeling of smart grid using a Software Engineering formalism. The starting point of this work it was that a smart home can be considered as an integrated system consists of various objects and system, that communicates and interact  with  each  other. The approach is base in Petri nets and from the premise that modeling of the smart home leads to clear understanding the overall behavior of the smart grid.

\section{Automated Verification Using Model Checking}
\label{sec:AutomatedVerification}
%--------------------------------------------------
%It is necessary to keep in mind that 
Validation is the process of determining whether a design meets the user requirements, whereas verification is the process of determining whether a design meets a set of requirements, specifications, and regulations~\cite{ClarkeHV18}. If the requirements, specifications, and regulations are given in a formal language, then it may be possible to automate the verification process, thus resulting in a process known as \textit{formal verification}. Verification may form part of a validation process.
%, but in general, validation cannot be formalized because it relates a system design to intent.  
While simulation and testing explore some of the possible behaviors and scenarios of the system, leaving open the question of whether the unexplored trajectories may contain a flaw, formal verification conducts an exhaustive exploration of all possible behaviors. Thus, when a design is pronounced correct by a formal verification method, it implies that all behaviors have been explored, and the questions of adequate coverage or a missed behavior becomes irrelevant \cite{Clarke2012}.
%Simulation may also be used for validation, but it is more problematic for verification.
%In order to use simulation for verification, it is necessary to ensure adequate coverage of operating conditions, scenarios, and system inputs. Testing can also be used for validation, however it is not feasible to cover all the aspects of the design space, and some times the real system must be used instead of a computing model.
 %
